---
title: "Data Analysis with Sparklyr"
description: |
  Data analysis with Sparklyr
author:
  - name: Andrea Spano
    url: andreaspano.github.io
    affiliation: Quantide
    affiliation_url: www.quantide.com 
date: "`r Sys.Date()`" 
output:
  distill::distill_article:
    self_contained: false
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Pkgs

Load required packages


```{r}
require(sparklyr)
require(dplyr)
require(readr)
require(ggplot2)
```

## Set up 

Spark environment setup 


```{r}
# global spark memory
Sys.setenv("SPARK_MEM" = "32g")
# Initialize configuration with defaults
config <- spark_config()
# Memory
config["sparklyr.shell.driver-memory"] <- "32g"
# Cores
config["sparklyr.connect.cores.local"] <- 6
```

## Connect 

Create local Spark context 

```{r context}
sc <- spark_connect(master = "local", version = "3.0", config = config)
```

## csv files

Load csv files

```{r read csv} 
system.time({
d_all <- spark_read_csv(sc , 
               name = 'd_all',
               path = "file:///data/airline/csv",
               header = T)
})
```

Save as parquet 

```{r write parquet}
system('rm -rf /data/airline/prq/')
system.time({
  spark_write_parquet(d_all, path = '/data/airline/prq/')
})
```

Disconnect from spark context and reload

```{r disconnect and reconnect}
spark_disconnect(sc)
sc <- spark_connect(master = "local", version = "3.0", config = config)
```


load parquet

```{r  read parquet}
system.time({
  d_all <- spark_read_parquet(sc , 
                          name = 'd_all',
                          path = "file:///data/airline/prq/", 
                          header = T)
  
})
```


## Check data size

```{r check data size}
system.time({
n <- d_all %>% 
  summarise(n = n()) %>% 
  collect()

print(n) 
})
```


##  prepare  model data  

Select variables of interests only 

```{r dplyr}
system.time({
d_model <- d_all %>%
  mutate(ArrDelay = as.numeric(ArrDelay) ,
         DepDelay = as.numeric(DepDelay) ,
         Distance = as.numeric(Distance)) %>% 
  filter(!is.na(ArrDelay) & !is.na(DepDelay) & !is.na(Distance)) %>%
  filter(DepDelay > -10 & DepDelay < 240) %>%
  filter(ArrDelay > -60 & ArrDelay < 360) %>% 
  mutate(Gain = DepDelay - ArrDelay) %>%
  select(Year, Month, ArrDelay, DepDelay, Distance, UniqueCarrier, Gain)
})
```

Check newdata dimension 

```{r count 2}
system.time({
n <- d_model %>% 
  summarise(n = n()) %>% 
  collect()
print(n)
})

```

import carriers data 

```{r import carriers}
d_carrier <- spark_read_csv(sc , 
                          name = 'd_carrier',
                          path = "file:///data/airline/csv/carriers.csv", 
                          header = T)
```

join airlines data with carriers data

```{r left join }
system.time({
d_model <- d_model %>% 
  left_join(d_carrier, by = c("UniqueCarrier" = "Code"))
})
```

split data into training and out of sample test 

```{r split}
system.time({
  d_model_2008 <- d_model %>% filter(Year == 2008)
  d_model <- d_model %>% filter(Year <= 2007)
})
```


check dimensions 

```{r check 3}
system.time({
n <- d_model %>%  summarise(n = n())
n2008 <- d_model_2008 %>%  summarise(n = n())
print(n)
print(n2008)
})


```
 
partition the data into training and validation sets

```{r partition }
system.time({
model_partition <- d_model %>% 
  sdf_random_split(d_trn = 0.8, d_tst = 0.2, seed = 5555)
})
```

## Modelling 

Fit a linear model

```{r model }
system.time({
  fm <-  ml_linear_regression(model_partition$d_trn, 
                            formula = Gain ~ Distance + DepDelay + UniqueCarrier)
})
```

Show summary

```{r sumnmary}
system.time({  
  summary(fm) 
})
```

Calculate average gains by predicted decile


```{r gain}
system.time({
model_deciles <- lapply(model_partition, 
                        function(x) {
                            ml_predict(fm, x) %>%
                              mutate(Decile = ntile(desc(prediction), 10)) %>%
                              group_by(Decile) %>%
                              summarize(Gain = mean(Gain)) %>%
                              select(Decile, Gain) %>%
                              collect()}
                        )
})
```


Create a summary dataset for plotting

```{r for plot }
d_decile <- bind_rows(
  as_tibble(model_deciles$d_trn) %>% mutate(partition = 'trn'),
  as_tibble(model_deciles$d_tst) %>% mutate(partition = 'tst'))
```
  

Plot average gains by predicted decile

```{r plot}
d_decile %>%
  ggplot(aes(factor(Decile), Gain, fill = partition)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  labs(title = 'Average gain by predicted decile', x = 'Decile', y = 'Minutes')
```

prediction time ahead

```{r pred}
system.time({
pred_2008 <- ml_predict(fm, d_model_2008) %>%
  group_by(Description) %>%
  summarize(Gain = mean(Gain), prediction = mean(prediction), freq = n()) %>%
  filter(freq > 10000) %>%
  collect()
})
```

Plot actual gains and predicted gains by airline carrier

```{r plot 2}
ggplot(pred_2008, aes(Gain, prediction)) + 
  geom_point(alpha = 0.75, color = 'red', shape = 3) +
  geom_abline(intercept = 0, slope = 1, alpha = 0.15, color = 'blue') +
  geom_text(aes(label = substr(Description, 1, 20)), size = 3, alpha = 0.75, vjust = -1) +
  labs(title='Average Gains Forecast', x = 'Actual', y = 'Predicted')
```


## Closing 

Disconnect from spark context

```{r disconnect}
spark_disconnect(sc)
```
