---
title: "Substitute"
description: |
  Understanding R function substitute 
author:
  - name: Andrea Spano
    url: www.quantide.com
    affiliation: Quantide
    affiliation_url: www.quantide.com
date: "`r Sys.Date()`"
output: distill::distill_article
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction 

The scope of this article is to clarify, at least for me, the behaviour of function `substitute()`. 

As a confusing example consider:

```{r}
x <- 0
substitute( x  )
```      

and 

```{r}
f <- function() {
  x <- 0
  substitute(x)
}  
f()
```      

As you can see function `substitute()` returns different result when called either from within the global environment or within a function environment

# Basic Substitute 

Assuming some prior knowledge of `environment` and `promise` objects you can analyse the documentation for `substitute()`. 

Documentation says:

Substitution takes place by examining each component of the parse tree as follows:

* If it is not a bound symbol in env, it is unchanged.
* If it is a promise object (i.e., a formal argument to a function) the expression slot of the promise replaces the symbol.
* If it is an ordinary variable, its value is substituted, unless env is .GlobalEnv in which case the symbol is left unchanged.

In practice, function `substitutes()` takes to arguments:

* expr 
* env

The return value of the function depends on the interaction between the two arguments.


We will take into account four cases corresponding to four types of objects to pass to argument `expr`

1. Non existing object 
2. variable
3. promise 
4. expression

For each object we will examine the interaction with the `env` argument

## Substitute non existing object

When symbol `x` does not exist within `env` Symbol `x` is returned


```{r}
env <- new.env()
substitute(expr = x , env = env)
```


## Substitute constant

When `x` is a variable in `env` and `env` is not the global environment, the value of the variable is returned.


```{r}
env <- new.env()
substitute(0, env = env)

substitute(0, env = globalenv())

```


## Substitute variable

When `x` is a variable in `env` and `env` is not the global environment, the value of the variable is returned.


```{r}
env <- new.env()
assign('x', 0, envir = env)
substitute(x, env = env)
```


When `x` is a variable  `globalenv`, the value of the variable the symbol `x` is returned

```{r}
assign('x', 0, envir = globalenv())
substitute(x, env = globalenv())
```



## Substitute promise

When `x` is a promise in `env`, the the expression slot of the promise is returned.


```{r}
env <- new.env()
delayedAssign('x', 0, assign.env = env)
substitute(x, env = env)
```


When the expression is a function formal, therefore a promise, the same rule applies: 

```{r}
f <- function(x) {
  substitute(x)
}
f(x = 0)
```  


## Substitute expression

<!---- ref: http://adv-r.had.co.nz/Computing-on-the-language.html --->

```{r}
x <- 1
exp <- quote(x+1)
substitute(exp, env = globalenv())

env <- new.env()
env$x <- 1
env$exp <- quote(x+1)
substitute(exp, env = env)



assign('x', 1+1 , envir = env)
env$x
```

```{r}
env <- new.env()
assign('x', 1+1 , envir = env)
env$x
```








## Recap 

In summary:

1. When the evaluation occurs outside the global environment, the value of the variable or the value slot of the promise is returned  

2. When evaluation occurs in the global environment the symbol is returned unchanged.

3. When the symbol does not exists in the evaluation environment, the symbol is returned unchanged.







