---
title: "Substitute"
description: |
  Understanding  function `substitute()` 
author:
  - name: Andrea Spano
    url: www.quantide.com
    affiliation: Quantide
    affiliation_url: www.quantide.com
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    self_contained: false
editor_options: 
  chunk_output_type: console
draft: false  
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction 

The scope of this article is to clarify, at least for me, the behaviour of function `substitute()`. 

As a confusing example consider:

```{r}
x <- 0
substitute( x  )
```      

and 

```{r}
f <- function() {
  x <- 0
  substitute(x)
}  
f()
```      

As you can see function `substitute()` returns different result when called either from within the global environment or within a function environment

# Basic Substitute 

Assuming some prior knowledge of `environment` and `promise` objects you can analyse the documentation for `substitute()`. 

Documentation says:

Substitution takes place by examining each component of the parse tree as follows:

* If it is not a bound symbol in env, it is unchanged.
* If it is a promise object (i.e., a formal argument to a function) the expression slot of the promise replaces the symbol.
* If it is an ordinary variable, its value is substituted, unless env is .GlobalEnv in which case the symbol is left unchanged.

In practice, function `substitutes()` takes to arguments:

* expr: any `R` expression
* env: either an environment or a list 

The return value of the function depends on the interaction between the two arguments.

## Test

We will take into account five cases corresponding to four types of objects to pass to argument `expr`

1. Non existing object 
1. Constant
1. variable
1. promise 
1. expression

For each object we will examine the interaction with the `env` argument

### Substitute non existing object

When symbol `x` does not exist within `env` Symbol `x` is returned


```{r}
substitute(expr = x , env = list())
```


### Substitute constant

When `x` is a constant either in `env` or `globalenv` the constant is returned.


```{r}
substitute(0, env = list())
substitute(0, env = globalenv())
```


### Substitute variable

When `x` is a variable in `env` and `env` is not the global environment, the value of the variable is returned.


```{r}
substitute(x, env = list ( x = 0 ))
```


When `x` is a variable  `globalenv`, the symbol `x` is returned

```{r}
x <- 0 
substitute(x, env = globalenv())
```



### Substitute promise

When `x` is a promise in `env`, the the expression slot of the promise is returned.

```{r}
env <- new.env()
delayedAssign('x', 0, assign.env = env)
substitute(x, env = env)
```

Note that argument `assign.env` of function `delayedAssign()` strictly require an environment.  

When the expression is a function formal, therefore a promise, the same rule applies: 

```{r}
f <- function(x) {
  substitute(x)
}
f(x = 0)
```  


When `x` is a promise in `globalenv`, the symbol slot of the promise is returned. Basically, as soon as the promise object is accessed the variable is evaluated and then passed to substitute 

```{r}
delayedAssign('x', 0, assign.env = globalenv())
substitute(x, env = globalenv())
```




### Substitute expression

<!---- ref: http://adv-r.had.co.nz/Computing-on-the-language.html --->

When exp is an `R` expression within `env` substitute returns the expression

```{r}
substitute(e, env = list( e = quote(x+1)))
```

When exp is an `R` expression within `globalenv` substitute returns the symbol associated to the expression 

```{r}
e <- quote(x+1)
substitute(expr = e, env = globalenv())
```


### Recap 

When function `substitute()` is used outside `globalenv`, it really substitutes 

```{r}
f <- function(){
  a <- 1
  b <- 2
  substitute(a+b+c)
}
f()
```


When function `substitute()` is used inside `globalenv`, no substitution occurs and the names associated to either the variable, the promise or the expression is returned unchanged. In practice, function `substitute()` within `globalenv()` works like function `quote()` 

```{r}
a <- 1
b <- 2
substitute(a+b+c)
```

## Working with substitute 

Because of its behaviour, developing and experimenting with function `substiute()` may be difficult.

In order to overcome this problem, Hadley in [adv-r](http://adv-r.had.co.nz/Computing-on-the-language.html)  proposes function `subs()` in package `pryr`:

```{r}
pryr::subs
```


Function `subs()` has the same behavior either inside or outside `globalenv`:

Within globalenv:

```{r}
x <- 1
pryr::subs(x+1)
```

Outside globalenv:

```{r}
f <- function(x) {
  pryr::subs(x+1)
}
f(1)
```




The inner part of function `subs()` relies on function `substitute_q`. 

A less formal, but more esplicative implementation of `substitute_q()` is 

```{r}
substitute_q <- function (x, env) {
    call <- substitute(substitute(y, env), list(y = x))
    eval(call)
}
```

When function `substitute_q()` is called fron `globalenv`, it returns the evaluated expression 

```{r}
substitute_q  (quote(a + 1), env = list ( a = 2)) 
```

as it does when called from any other environment

```{r}
f <- function(x)  {
  a <- 2
  substitute_q(x, env = environment())
}

f(quote(a+1))
```






