---
title: "Interactive Plot_ly"
description: |
  The idea is about using package `plotly` for producing  interactive plots for a Shiny application.  
  Official documentation about `plotly` is superb. 
  Nevertheless, as this graphics grammar, within the `R` community is not as spread as `ggplot` is, few topics require a bit of extra research.
author:
  - name: Andrea Spano
    url: andreaspano.github.io
    affiliation: Quantide
    affiliation_url: www.quantide.com 
date: "`r Sys.Date()`" 
output:
  distill::distill_article:
    self_contained: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Intro

Given 

*  a data set with four columns: `x, y , w, z` 
*  a plot `x` and `y` onto `pl1`
*  plot `w` and `z` onto `pl2`
* `pl` as a subplot of `pl1` and `pl2`

I would like to:

* be able to select a set of contiguos points in `pl1` 
* observe the selected points marked with new color and larger size
* observe the same effect on corresponding points on pl2

I will be using the classic `iris` data for this example.


## Step 1: 

Load the required packages

```{r}
require(plotly)
require(dplyr)
```


## Step 2: 

Add a unique id to data 


```{r}
data <- iris %>% mutate ( id = 1:n())
```


## Step 3: 

Transform  data into a `ShareData` object

```{r}
data <- data %>% highlight_key(~id) 
```


## Step 4:

Build your plots as you would normally do

``` {r}
pl1 <- plot_ly(data, 
               x = ~ Sepal.Length, y = ~Sepal.Width, 
               type = 'scatter', mode = 'markers', 
               marker = list ( color = 'green')) %>% 
  layout(showlegend=F)
               
pl2 <- plot_ly(data, 
               x = ~Petal.Length, y = ~Petal.Width, 
               type = 'scatter', mode = 'markers',
               marker = list ( color = 'blue')) %>% 
  layout(showlegend=F)
```

## Step 5: 

Bind your plots into a single `subplot`
      
```{r}      
pl <- subplot(pl1, pl2, nrows = 2)
```

## Step 6:

Link the plots

```{r}
pl <- pl %>% 
  highlight(on = "plotly_select", off = 'plotly_doubleclick', 
            opacityDim = .5, 
            selected = attrs_selected(marker = list ( size = 20, color = 'orange')))
```
  
## Step 7: 

Display results

```{r}
pl
```








