---
title: "Run an R script as a background job"
description: |
  How to run an R script from within R in background mode
author:
  - name: Andrea Spano
    url: andreaspano.github.io
    affiliation: Quantide
    affiliation_url: www.quantide.com
date: "`r Sys.Date()`"
bibliography: 
  biblio.bib
output:
  distill::distill_article:
    self_contained: false
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

When working with R you may want to run a script that require quite a long time to complete. 

As a basic non sense example consider:

```{r, eval = FALSE}
n <- 1200
y <- numeric(n)
for ( i in seq_len(n)){
  y[[i]] <- i 
  Sys.sleep(3)
}
sum(y)
```

This script, say `wait-for-me.R` when executed takes about one hour. 

If executed directly from the R console it will keep your R session busy for one hour preventing you to keep working on you current R session. 

In order to avoid this waste of time you can use the function `jobRunScript()` from package `rstudioapi` [@rstudioapi]

```{r,  eval = FALSE}
require(rstudioapi)
jobRunScript(path = './wait-for-me.R', 
             name = 'wait for me',
             encoding = "unknown", 
             workingDir = NULL,
             importEnv = FALSE, 
             exportEnv = "R_GlobalEnv")
```

This command explains to R to run your script and export the objects created within this environment  


3600/3
